Off the beaten path: \\ ...

### 2018-04-28 paper/gd1.tex
Tweaks

The most prominent and longest (in angular extent) thin stream is the GD-1
stream, named after its discoverers (\citealt{Grillmair:2006}).
GD-1 was found using a matched filter on photometric data from the SDSS and
spans over $60^\circ$ in Heliocentric sky coordinates, partially because of its
relative closeness to the sun ($\approx 7$--$10~\textrm{kpc}$).
Despite its proximity, no remnant progenitor cluster or stellar system has been
found associated with the stream, but its metallicity ($\feh \approx -1.4$;
\citealt{Koposov:2010}) and estimated stellar mass ($M \approx 2 \times
10^4~\msun$; \citealt{Koposov:2010}) are consistent with being a disrupted
globular cluster.

The long physical length of the GD-1 stream ($\sim 15~\kpc$) and location in the
Galactic halo (pericentric distance $r_\textrm{peri} \sim 13~\kpc$) make it an
ideal object for constraining the gravitational potential of dark matter in the
inner Milky Way.
Indeed, both from fitting orbits to binned phase-space measurements along the
GD-1 stream (\citealt{Koposov:2010}) and from modeling the stream track in
phase-space coordinates (\citealt{Bovy:2016}), it has been shown that the dark
matter distribution within its orbit is consistent with or nearly spherical.

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
By assuming an age [...], \citet{Erkal:2016} predict that XX (~1 gap??).

From SDSS photometry, hints of further gaps and under-densities have been
identified in the GD-1 stream, but with low significance
(\citealt{Carlberg:2013}).
More recently, deep photometry of GD-1 from CFHT/Megacam has revealed
interesting small-scale ``wiggles'' and significant density variations along a $\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}):
[summarize their conclusions].

>>>>

The most prominent and longest (in angular extent) thin stream is the GD-1
stream (named after its discoverers; \citealt{Grillmair:2006}).
GD-1 was found using a matched filter on photometric data from the SDSS and
was found to span $\approx 60^\circ$ in Heliocentric sky coordinates, partially
because of its relative proximity to the sun ($\approx 7$--$10~\textrm{kpc}$).
Despite being close and relatively high surface-density, no remnant progenitor
cluster or stellar system has been found associated with the stream, but its
width,  metallicity ($\feh \approx -1.4$; \citealt{Koposov:2010}) and estimated
stellar mass ($M \approx 2 \times 10^4~\msun$; \citealt{Koposov:2010}) are
consistent with being a disrupted globular cluster.

The long physical length of the GD-1 stream ($\sim 15~\kpc$) and location in the
Galactic halo (pericentric distance $r_\textrm{peri} \sim 13~\kpc$) make it an
ideal object for constraining the gravitational potential of dark matter in the
inner Milky Way.
Indeed, both from fitting orbits to binned phase-space measurements along the
GD-1 stream (\citealt{Koposov:2010}) and from modeling the stream track in
phase-space coordinates (\citealt{Bovy:2016}), it has been used to show that the
dark matter distribution within its orbit is consistent with being close to
spherical.

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
By estimating a stream age of $\approx 4~\textrm{Gyr}$ assuming a progenitor
mass of $10^5~\msun$, and by taking a subhalo mass function and density from
XXX, \citet{Erkal:2016} predict that GD-1 could have up to one significant, wide
($\approx 5^\circ$--$7^\circ$) gap caused by an interaction with a
$10^6$--$10^7~\msun$ subhalo.

TODO: hints of density variations from initial work Koposov:2010, later confirmed by density modeling, but all low significance
(\citealt{Carlberg:2013}).
More recently, deep photometry of GD-1 from CFHT/Megacam has revealed
interesting small-scale ``wiggles'' and significant density variations along a $\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}):
[summarize their conclusions].
